---
title: "Norma Evaluation Diffing"
output: html_document
date: "`r Sys.Date()`"
params:
    datafile: "measurements.csv"
---

```{r setup, echo=FALSE, include=FALSE}
library(tidyverse)

figure_dimensions <- c(10, 4)

# load monitoring data
all_data <- read_csv(params$datafile)

# comput the start time of each run
start_times <- all_data %>%
    dplyr::filter(metric == "SentTransactions") %>%  # filter metric
    select(run, time) %>%
    group_by(run) %>%
    summarise(start = min(time))

# join the start times into the monitoring data
all_data_with_start_time <- inner_join(all_data, start_times, by = "run")
```

This report compares the execution of multiple Norma evaluations.

## Application Metrics

### Transactions Send vs. Received

The following chart shows the total number of sent and received transactions in each of the scenarios.

```{r send_and_received_transactions_per_run, echo=FALSE, message=FALSE, fig.dim = figure_dimensions}
# compute the send rate for each account
sent_rate <- all_data_with_start_time %>%
    dplyr::filter(metric == "SentTransactions") %>%   # filter metric
    select(run, time, app, workers, value, start) %>% # filter relevant columns
    mutate(value = as.numeric(value)) %>%             # convert value to integer
    group_by(run, app, workers) %>%                   # one group per account
    arrange(app, workers, time) %>%                   # sort by time in groups
    reframe(
        run = run,                                    # preserve the run
        time = time,                                  # preserve the time
        start = start,                                # preserve the start time
        rate = value - lag(value, default = 0)        # change since last
    )

# compute the accumulated number of send transactions per run
sent <- sent_rate %>%
    group_by(run) %>%                                # group by run
    arrange(time) %>%                                # sort by time in groups
    reframe(time = time, start = start, total = cumsum(rate)) %>%   # compute prefix sum
    mutate(simtime = as.numeric(time) - as.numeric(start))       # compute relative time

# compute the number of received transactions
received <- all_data_with_start_time %>%
    dplyr::filter(metric == "ReceivedTransactions") %>%  # filter metric
    mutate(value = as.numeric(value)) %>%           # convert value to int
    mutate(simtime = as.numeric(time) - as.numeric(start))      # compute relative time

ggplot() +
    geom_line(data=sent, aes(x=simtime/1e9, y=total, group=run, colour = factor(paste(run, " sent")))) +
    geom_line(data=received, aes(x=simtime/1e9, y=value, group=run, colour = factor(paste(run, " received")))) +
    ggtitle("Sent and Received Transactions") +        # chart title
    xlab("Simulation Time [s]") +                      # x-axis title
    ylab("Total Number of Transactions") +             # y-axis title
    labs(colour="Runs") +                              # legend title
    theme(plot.title = element_text(hjust = 0.5))      # center title
```

The following chart compares the througput of the runs over time.

```{r throughput_over_time, echo=FALSE, message=FALSE, fig.dim = figure_dimensions}
# compute the number of received transactions
received <- all_data_with_start_time %>%
    dplyr::filter(metric == "ReceivedTransactions") %>%  # filter metric
    mutate(value = as.numeric(value)) %>%           # convert value to int
    mutate(time = as.numeric(time)) %>%             # convert time to int
    mutate(start = as.numeric(start)) %>%           # convert start to int
    mutate(simtime = time - start)                  # compute relative time

# compute the receive rate over time
received_rate <- received %>%
    group_by(run) %>%
    arrange(time) %>%
    reframe(
        simtime = simtime,
        rate = (value - lag(value, default = 0)) / ((time - lag(time)) / 1e9)
    ) %>%
    drop_na(rate)

ggplot() +
    geom_line(data=received_rate, aes(x=simtime/1e9, y=rate, group=run, colour = factor(run))) +
    ggtitle("Received Transactions") +                 # chart title
    xlab("Simulation Time [s]") +                      # x-axis title
    ylab("Tx/s") +                                     # y-axis title
    labs(colour = "Runs") +                            # legend title
    theme(plot.title = element_text(hjust = 0.5))      # center title
```

```{r throughput_over_time_smoothed, echo=FALSE, message=FALSE, fig.dim = figure_dimensions}
ggplot() +
    geom_smooth(data=received_rate, aes(x=simtime/1e9, y=rate, group=run, colour = factor(run))) +
    ggtitle("Received Transactions (Smoothed)") +      # chart title
    xlab("Simulation Time [s]") +                      # x-axis title
    ylab("Tx/s") +                                     # y-axis title
    labs(colour = "Runs") +                            # legend title
    theme(plot.title = element_text(hjust = 0.5))      # center title
```

## Node Metrics

### Block Processing Time

The following chart shows the time required by each node for processing each individual block.

```{r block_processing_time, echo=FALSE, message=FALSE, fig.dim = figure_dimensions}
data <- all_data %>%
    dplyr::filter(metric == "BlockEventAndTxsProcessingTime") %>%
    mutate(value = as.numeric(value)) %>%
    mutate(label = paste(run, node))

ggplot(data = data) +
    geom_point(aes(x=block, y=value/1e6, group=label, colour = factor(label))) +
    ggtitle("Block Processing Time") +            # chart title
    xlab("Block Height") +                        # x-axis title
    ylab("Block Processing Time [ms]") +          # y-axis title
    labs(colour = "Nodes") +                      # legend title
    theme(plot.title = element_text(hjust = 0.5)) # center title
```

While the individual sample points may provide the highest level of detail, the smoothed out trend lines of the following plot may provide a cleaner comparison between the processing times of individual nodes.

```{r block_processing_time_smooth, echo=FALSE, message=FALSE, fig.dim = figure_dimensions}
ggplot(data = data) +
    geom_smooth(aes(x=block, y=value/1.e6, group=label, colour = factor(label))) +
    ggtitle("Block Processing Time") +            # chart title
    xlab("Block Height") +                        # x-axis title
    ylab("Block Processing Time [ms]") +          # y-axis title
    labs(colour = "Nodes") +                      # legend title
    theme(plot.title = element_text(hjust = 0.5)) # center title
```
