---
title: "Norma Evaluation Summary"
output: html_document
date: "`r Sys.Date()`"
params:
    datafile: "measurements.csv"
---

```{r setup, echo=FALSE, include=FALSE}
library(tidyverse)
library(lubridate)
library(slider)

figure_dimensions <- c(10, 4)

# load monitoring data
all_data <- read_csv(params$datafile)
```

This report summarizes the monitoring results of an execution of a Norma scenario.

## Block Metrics

This section covers per-Block metrics which are consistent throughout the network.

### Gas per Block
The following chart shows the Gas spent per block and a trend line showing long-term trends.


```{r gas_spent_per_block, echo=FALSE, message=FALSE, fig.dim = figure_dimensions}
data <- all_data %>%
    dplyr::filter(metric == "BlockGasUsed") %>%    # filter metric of interrest
    mutate(value = as.numeric(value))              # convert value to int

ggplot(data=data) +
    geom_point(aes(x=block, y=value)) +               # the data points
    geom_smooth(aes(x=block, y=value), se=FALSE) +    # a trend line
    ggtitle("Gas per Block") +                        # chart title
    xlab("Block Height") +                            # x-axis title
    ylab("Gas spent") +                               # y-axis title
    theme(plot.title = element_text(hjust = 0.5))     # center title
```


### Transactions per Block
The following chart shows the number of Transactions per block and a trend line showing long-term trends.

```{r transactions_per_block, echo=FALSE, message=FALSE, fig.dim = figure_dimensions}
data <- all_data %>%
    dplyr::filter(metric == "BlockNumberOfTransactions") %>%   # filter metric
    mutate(value = as.numeric(value))                          # convert to int

ggplot(data=data) +
    geom_point(aes(x=block, y=value)) +               # the data points
    geom_smooth(aes(x=block, y=value), se=FALSE) +    # a trend line
    ggtitle("Transactions per Block") +               # chart title
    xlab("Block Height") +                            # x-axis title
    ylab("Number of Transactions") +                  # y-axis title
    theme(plot.title = element_text(hjust = 0.5))     # center title
```


### Block Delay
The following chart shows the time between the earliest reported completion time of block `n-1` and earliest reported completion time block `n` for block `n`.

```{r block_delay_time, echo=FALSE, message=FALSE, fig.dim = figure_dimensions}
data <- all_data %>%
    # filter desired metric and columns
    dplyr::filter(metric == "BlockCompletionTime") %>%
    select(block, value) %>%
    # convert values into numbers
    mutate(value = as.numeric(value)) %>%
    # find earliest time for each block
    group_by(block) %>%
    summarise(value = min(value)) %>%
    # compute delta between consecutive blocks
    arrange(block) %>%
    mutate(delta = value - lag(value)) %>%
    drop_na(delta)

ggplot(data = data) +
    # a horizontal line marking the 1s target
    geom_hline(yintercept = 1, linetype = "dashed", color = "red") +
    # the individual block delays
    geom_point(aes(x = block, y = delta / 1e9)) +
    # a interpolation line
    geom_smooth(aes(x = block, y = delta / 1e9), se = FALSE) +
    ggtitle("Block Delay") +                           # chart title
    xlab("Block Height") +                             # x-axis title
    ylab("Delay [s]") +                                # y-axis title
    theme(plot.title = element_text(hjust = 0.5))      # center title
```


## Network Metrics

This section covers metrics that are network wide properties.

### Network Incoming and Committed Transaction Rates

The following chart shows the total rate of incoming and committed transactions smoothed over 5 second intervals.

```{r throughput_over_time, echo=FALSE, message=FALSE, fig.dim = figure_dimensions}
# compute the rate at which transactions have been sent to the network
sent_rate <- all_data %>%
    # filter for rows and columns needed
    dplyr::filter(metric == "SentTransactions") %>%
    select(time, app, workers, value) %>%
    # convert values to numerical values
    mutate(value = as.numeric(value)) %>%
    mutate(time = as.numeric(time)) %>%
    # convert per-account absolute values to increment values
    group_by(app, workers) %>%
    arrange(time) %>%
    reframe(
        time = time,
        increment = value - lag(value, default = 0)
    ) %>%
    # compute average rate over dynamic sized windows (each 5 seconds)
    arrange(time) %>%
    slide_period_dfr(
        as.POSIXct(.$time/1e9,origin="1970-01-01"),
        "second",
        .every = 5,
        function(data) {
            summarize(
                data,
                # compute the rate over time window
                rate = sum(increment) / 5,
                # associate rate to mid-point time
                time = (max(time)+min(time))/2
            )
        }
    )

if (nrow(sent_rate) > 0) {
    sent_rate <- sent_rate %>%
        drop_na(rate) %>%
        # convert time from UNIX to DateTime type
        mutate(date = as_datetime(time/1e9))
} else {
    sent_rate <- sent_rate %>%
        mutate(rate = 0) %>%
        mutate(date = as_datetime(0))
}


# compute the number of received (=committed) transactions
received_rate <- all_data %>%
    # filter rows and columns
    dplyr::filter(metric == "ReceivedTransactions") %>%
    select(time, value) %>%
    # convert value and time to numerical values
    mutate(value = as.numeric(value)) %>%
    mutate(time = as.numeric(time)) %>%
    # compute the individual increments
    arrange(time) %>%
    reframe(
        time = time,
        increment = value - lag(value, default=0),
    ) %>%
    # compute average rate over dynamic sized windows (each 5 seconds)
    slide_period_dfr(
        as.POSIXct(.$time/1e9,origin="1970-01-01"),
        "second",
        .every = 5,
        function(data) {
            summarize(
                data,
                # compute the rate over time window
                rate = sum(increment) / 5,
                # associate rate to mid-point time
                time = (max(time)+min(time))/2
            )
        }
    )

if (nrow(sent_rate) > 0) {
    received_rate <- received_rate %>%
        drop_na(rate) %>%
        # convert time from UNIX to DateTime type
        mutate(date = as_datetime(time/1e9))
} else {
    received_rate <- received_rate %>%
        mutate(rate = 0) %>%
        mutate(date = as_datetime(0))
}

ggplot() +
    geom_line(data=sent_rate, aes(x=date, y=rate, color = "incomming [5s window]")) +
    #geom_smooth(data=sent_rate, aes(x=date, y=rate)) +
    geom_line(data=received_rate, aes(x=date, y=rate, color = "committed [5s window]")) +
    #geom_smooth(data=received_rate, aes(x=date, y=rate)) +
    ggtitle("Incomming and Committed Transaction Rates") +
    xlab("Time") +
    ylab("Tx/s") +
    theme(plot.title = element_text(hjust = 0.5))
```


### Number of Nodes in the Network

The number of nodes in the network over time.

```{r nodes_in_network, echo=FALSE, message=FALSE, fig.dim = figure_dimensions}
data <- all_data %>%
    dplyr::filter(metric == "NumberOfNodes") %>%      # filter metric
    mutate(date = as_datetime(as.numeric(time)/1e9)) %>%  # UNIX time to date
    mutate(value = as.numeric(value))                 # convert value to int

ggplot(data=data) +
    geom_point(aes(x=date, y=value)) +               # the data points
    ggtitle("Nodes in Network") +                    # chart title
    xlab("Time") +                                   # x-axis title
    ylab("Number of Nodes") +                        # y-axis title
    theme(plot.title = element_text(hjust = 0.5)) +  # center title
    scale_x_datetime(date_labels = "%c")             # format date labels
```


## Node Metrics

This section covers metrics that are node specific.

### Block Processing Time

The following chart shows the time required by each node for processing each individual block.

```{r block_processing_time, echo=FALSE, message=FALSE, fig.dim = figure_dimensions}
data <- all_data %>%
    dplyr::filter(metric == "BlockEventAndTxsProcessingTime") %>%
    mutate(value = as.numeric(value))

ggplot(data=data) +
    geom_point(aes(x=block, y=value/1e6, group=node, colour = factor(node))) +
    ggtitle("Block Processing Time") +            # chart title
    xlab("Block Height") +                        # x-axis title
    ylab("Block Processing Time [ms]") +          # y-axis title
    labs(colour="Nodes") +                        # legend title
    theme(plot.title = element_text(hjust = 0.5)) # center title
```

While the individual sample points may provide the highest level of detail, the smoothed out trend lines of the following plot may provide a cleaner comparison between the processing times of individual nodes.

```{r block_processing_time_smooth, echo=FALSE, message=FALSE, fig.dim = figure_dimensions}
ggplot(data=data) +
    geom_smooth(aes(x=block, y=value/1.e6, group=node, colour = factor(node))) +
    ggtitle("Block Processing Time") +            # chart title
    xlab("Block Height") +                        # x-axis title
    ylab("Block Processing Time [ms]") +          # y-axis title
    labs(colour="Nodes") +                        # legend title
    theme(plot.title = element_text(hjust = 0.5)) # center title
```


### Block Height per Node

The following char shows the time at which nodes have reached a given block height.

```{r block_height_per_node, echo=FALSE, message=FALSE, fig.dim = figure_dimensions}
data <- all_data %>%
    dplyr::filter(metric == "NodeBlockHeight") %>%  # filter metric of interrest
    mutate(date = as_datetime(as.numeric(time)/1e9)) %>%   # UNIX time to date
    mutate(value = as.numeric(value))               # convert value to int

ggplot(data=data) +
    geom_line(aes(x=date, y=value, group=node, colour = factor(node))) +
    ggtitle("Block Height over Time") +                # chart title
    xlab("Time") +                                     # x-axis title
    ylab("Block Height") +                             # y-axis title
    labs(colour="Nodes") +                             # legend title
    theme(plot.title = element_text(hjust = 0.5)) +    # center title
    scale_x_datetime(date_labels = "%c")               # format date labels
```

## Application Metrics

### Transactions Send vs. Received

The following chart shows the total number of sent and received transactions.

```{r send_and_received_transactions_per_app, echo=FALSE, message=FALSE, fig.dim = figure_dimensions}
# compute the send rate for each account
sent_rate <- all_data %>%
    dplyr::filter(metric == "SentTransactions") %>%  # filter metric
    select(time, app, workers,value) %>%             # filter relevant columns
    mutate(value = as.numeric(value)) %>%            # convert value to integer
    group_by(app, workers) %>%                       # one group per account
    arrange(app, workers, time) %>%                  # sort by time in groups
    reframe(
        time = time,                                 # preserve the time
        rate = value - lag(value, default = 0)       # change since last
    )

# compute the total number of send transactions per application
sent <- sent_rate %>%
    group_by(app) %>%                                # group by application
    arrange(app, time) %>%                           # sort by time in groups
    reframe(time = time, total = cumsum(rate)) %>%   # compute prefix sum
    mutate(date = as_datetime(as.numeric(time)/1e9)) # convert time

# compute the number of received transactions
received <- all_data %>%
    dplyr::filter(metric == "ReceivedTransactions") %>%  # filter metric
    mutate(date = as_datetime(as.numeric(time)/1e9)) %>%   # UNIX time to date
    mutate(value = as.numeric(value))               # convert value to int

ggplot() +
    geom_line(data=sent, aes(x=date, y=total, group=app, colour = factor(paste(app, " sent")))) +
    geom_line(data=received, aes(x=date, y=value, group=app, colour = factor(paste(app, " received")))) +
    ggtitle("Sent and Received Transactions per App") +         # chart title
    xlab("Time") +                                     # x-axis title
    ylab("Total Number of Transactions") +             # y-axis title
    labs(colour="Apps") +                              # legend title
    theme(plot.title = element_text(hjust = 0.5)) +    # center title
    scale_x_datetime(date_labels = "%c")               # format date labels
```


The following chart shows the cummulative number of transactions send per account  (legend omitted for clarity).

```{r sent_transactions_per_app, echo=FALSE, message=FALSE, fig.dim = figure_dimensions}
data <- all_data %>%
    dplyr::filter(metric == "SentTransactions") %>%  # filter metric
    mutate(date = as_datetime(as.numeric(time) / 1e9)) %>%   # UNIX time to date
    mutate(value = as.numeric(value)) %>%              # convert value to int
    mutate(account = paste(app, "_", workers))         # add an user name

ggplot(data = data) +
    geom_line(
        aes(x = date, y = value, group = account, colour = factor(account)),
        show.legend = FALSE
    ) +
    ggtitle("Sent Transactions per User") +            # chart title
    xlab("Time") +                                     # x-axis title
    ylab("Total Number of Transactions") +             # y-axis title
    labs(colour = "User") +                            # legend title
    theme(plot.title = element_text(hjust = 0.5)) +    # center title
    scale_x_datetime(date_labels = "%c")               # format date labels
```

The following chart shows the send-rate of individual accounts (legend omitted for clarity).

```{r send_rate_per_account, echo=FALSE, message=FALSE, fig.dim = figure_dimensions}
sent_rate <- sent_rate %>%
    mutate(date = as_datetime(as.numeric(time) / 1e9)) %>%
    mutate(account = paste(app, "_", workers))         # add an user name

ggplot(data = sent_rate) +
    geom_line(
        aes(x = date, y = rate, group = account, colour = factor(account)),
        show.legend = FALSE
    ) +
    ggtitle("Sent Rate of Transactions per User") +    # chart title
    xlab("Time") +                                     # x-axis title
    ylab("Rate [Tx/s]") +                              # y-axis title
    labs(colour = "User") +                            # legend title
    theme(plot.title = element_text(hjust = 0.5)) +    # center title
    scale_x_datetime(date_labels = "%c")               # format date labels
```
