---
title: "Norma Evaluation Summary"
output: html_document
date: "`r Sys.Date()`"
params:
    datafile: "measurements.csv"
---

```{r setup, echo=FALSE, include=FALSE}
library(tidyverse)
library(lubridate)

figure_dimensions = c(10, 4)

# load monitoring data
all_data <- read_csv(params$datafile)
```

This report summarizes the monitoring results of an execution of a Norma scenario.

## Block Metrics

This section covers per-Block metrics which are consistent throughout the network.

### Gas per Block
The following chart shows the Gas spent per block and a trend line showing long-term trends.


```{r gas_spent_per_block, echo=FALSE, message=FALSE, fig.dim = figure_dimensions}
data <- all_data %>%
    dplyr::filter(metric == "BlockGasUsed") %>%    # filter metric of interrest
    mutate(value = as.numeric(value))              # convert value into numeric values

ggplot(data=data) +
    geom_point(aes(x=block, y=value)) +               # the data points
    geom_smooth(aes(x=block, y=value), se=FALSE) +    # a smoothed out trend line
    ggtitle("Gas per Block") +                        # chart title
    xlab("Block Height") +                            # x-axis title
    ylab("Gas spent") +                               # y-axis title
    theme(plot.title = element_text(hjust = 0.5))     # center title
```


### Transactions per Block
The following chart shows the number of Transactions per block and a trend line showing long-term trends.

```{r transactions_per_block, echo=FALSE, message=FALSE, fig.dim = figure_dimensions}
data <- all_data %>%
    dplyr::filter(metric == "BlockNumberOfTransactions") %>%   # filter metric of interrest
    mutate(value = as.numeric(value))                          # convert value into numeric values

ggplot(data=data) +
    geom_point(aes(x=block, y=value)) +               # the data points
    geom_smooth(aes(x=block, y=value), se=FALSE) +    # a smoothed out trend line
    ggtitle("Transactions per Block") +               # chart title
    xlab("Block Height") +                            # x-axis title
    ylab("Number of Transactions") +                  # y-axis title
    theme(plot.title = element_text(hjust = 0.5))     # center title
```


## Network Metrics

This section covers metrics that are network wide properties.

### Number of Nodes in the Network

The number of nodes in the network over time.

```{r nodes_in_network, echo=FALSE, message=FALSE, fig.dim = figure_dimensions}
data <- all_data %>%
    dplyr::filter(metric == "NumberOfNodes") %>%      # filter metric of interrest
    mutate(date = as_datetime(as.numeric(time)/1e9)) %>%          # convert UNIX time into a date
    mutate(value = as.numeric(value))                 # convert value into numeric values

ggplot(data=data) +
    geom_point(aes(x=date, y=value)) +               # the data points
    ggtitle("Nodes in Network") +                    # chart title
    xlab("Time") +                                   # x-axis title
    ylab("Number of Nodes") +                        # y-axis title
    theme(plot.title = element_text(hjust = 0.5)) +  # center title
    scale_x_datetime(date_labels = "%c")             # format of the X axis labels to print date and time
```


## Node Metrics

This section covers metrics that are node specific.

### Block Processing Time

The following chart shows the time required by each node for processing each individual block.

```{r block_processing_time, echo=FALSE, message=FALSE, fig.dim = figure_dimensions}
data <- all_data %>%
    dplyr::filter(metric == "BlockEventAndTxsProcessingTime") %>%  # filter metric of interrest
    mutate(value = as.numeric(value))                              # convert value into numeric values

ggplot(data=data) +
    geom_point(aes(x=block, y=value/1e6, group=node, colour = factor(node))) +
    ggtitle("Block Processing Time") +            # chart title
    xlab("Block Height") +                        # x-axis title
    ylab("Block Processing Time [ms]") +          # y-axis title
    labs(colour="Nodes") +                        # legend title
    theme(plot.title = element_text(hjust = 0.5)) # center title
```

While the individual sample points may provide the highest level of detail, the smoothed out trend lines of the following plot may provide a cleaner comparison between the processing times of individual nodes.

```{r block_processing_time_smooth, echo=FALSE, message=FALSE, fig.dim = figure_dimensions}
ggplot(data=data) +
    geom_smooth(aes(x=block, y=value/1.e6, group=node, colour = factor(node))) +
    ggtitle("Block Processing Time") +            # chart title
    xlab("Block Height") +                        # x-axis title
    ylab("Block Processing Time [ms]") +          # y-axis title
    labs(colour="Nodes") +                        # legend title
    theme(plot.title = element_text(hjust = 0.5)) # center title
```


### Block Height per Node

The following char shows the time at which nodes have reached a given block height.

```{r block_height_per_node, echo=FALSE, message=FALSE, fig.dim = figure_dimensions}
data <- all_data %>%
    dplyr::filter(metric == "NodeBlockHeight") %>%  # filter metric of interrest
    mutate(date = as_datetime(as.numeric(time)/1e9)) %>%          # convert UNIX time into a date
    mutate(value = as.numeric(value))               # convert value into numeric values

ggplot(data=data) +
    geom_line(aes(x=date, y=value, group=node, colour = factor(node))) +
    ggtitle("Block Height over Time") +                # chart title
    xlab("Time") +                                     # x-axis title
    ylab("Block Height") +                             # y-axis title
    labs(colour="Nodes") +                             # legend title
    theme(plot.title = element_text(hjust = 0.5)) +    # center title
    scale_x_datetime(date_labels = "%c")               # format of the X axis labels to print date and time
```
